{\includegraphics[width=3cm, height=2.5cm]{Cran.jpg}} \\ \includegraphics[width=7cm, height=1cm]{TB.jpg} \\ {Technical Documentation for TersectBrowser+}

### 2025-05-02 TechDoc/TechDoc.tex
Edits to genome browser & gene models section

The function \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+. If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the constant \verb +JbrowseWithAccessionName+ is returned (See below parameters). If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated. If either of them are empty, a container with the phrase "Loading…" is returned. If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+. This takes as arguments the following variables: 
\begin{itemize}
\item \verb +assembly+: the assembly track 
\item \verb +tracks+: the data tracks to display 
\item \verb +configuration+: the config variable where the theme is defined 
\item \verb +defaultSession+: an object of type LinearGenomeView, which defines the initial state. This object includes the following configurations: 
\begin{itemize}
\item \verb +bpPerPx+: the base pairs per pixel, which defines the bin size 
\item \verb +assemblyName+: the corresponding assembly for the track 
\item \verb +start+: the track start position 
\item \verb +end+: the track end position 
\item \verb +refName+: the genomic coordinates of the viewing window 
\end{itemize}
\end{itemize}

When no accession is selected in the Canvas interface, the first three tracks defined in \verb +tracks.ts+ are added to the \verb +viewState+. The dynamic left-hand offset is set by \verb +horizontalScroll()+.

@@

\paragraph{Offset}
The dynamic offset is synced in a similar fashion to zoom, bin size, chromosome selection, and interval. The observable \verb +offsetCanvas+ is defined as a component of the \verb +PlotStateService+ class. The public variable \verb +offsetCanvasSource+ is defined as an instance of the \verb +BehaviourSubject+ class and holds all recorded values of the canvas offset. In the class constructor, \verb +offsetCanvas+ is initialised to continuously hold the latest value from \verb +offsetCanvasSource+. The canvas offset is set in the \verb +TreePlotComponent+ class, and is passed to \verb +offsetCanvasSource+ when the tree is redrawn.

Inside \verb +tersect-browser.component.ts+, the variable \verb +offsetCanvas+ is defined, and the subscription \verb +offsetCanvasSub+ listens to the  \verb +offsetCanvas+ observable and assigns the latest value to \verb +offsetCanvas+. Inside \verb +tersect-browser.component.html+, the \verb +offsetCanvas+ is passed to \verb +JbrowseWrapper+ as the prop \verb +offsetCanvas+. Inside \verb +JbrowseWrapper+, the variable defines the extent of the \verb +horizontalScroll+ using the following logic: $horizontalScroll(-(location.offsetCanvas - 4))$.

@@

When a new dataset is added, the custom addition script automatically updates the names of tracks and assembly in the relevant files within the genome-browser extension. This flexibility allows the Variant Browser extension to be accessible by plant breeders working on many different datasets. TersectBrowser was initially validated with a human genome dataset, so it was important to maintain this flexibility of input in the new extensions.

@@

\subsubsection{User Interface}
\paragraph{Feature Description}
\begin{itemize}
    \item The gene models track is displayed in the main JBrowse window. 
%    \item Gene model annotations are based on the official \hl{tomato genome annotation v2.4}: \href{https://phytozome-next.jgi.doe.gov/info/Slycopersicum_ITAG2_4#:~:text=Genome%20Overview,genome%2Cwith%2034%2C725%20gene%20models}{Slycopersicum ITAG annotation v2.4}
% The above is tomato-specific. Not sure where it fits. Probably doesn't fit anywhere - included it in notes so that we knew where the tomato gene models came from
    \item In the popup window, the gene models track is displayed as the second track, underneath the reference sequence track and above the variant track.
\end{itemize}

>>>>

The method \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+. If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the \verb +JbrowseWithAccessionName+ is returned (See below parameters). If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated. If either of them are empty, a container with the phrase "Loading…" is returned. If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+. This takes as arguments the following variables: 
\begin{itemize}
\item \verb +assembly+: the assembly track 
\item \verb +tracks+: the data tracks to display 
\item \verb +configuration+: the config variable where the theme is defined 
\item \verb +defaultSession+: an object of type LinearGenomeView, which defines the initial state. This object includes the following configurations: 
\begin{itemize}
\item \verb +bpPerPx+: the base pairs per pixel, which defines the bin size 
\item \verb +assemblyName+: the corresponding assembly for the track 
\item \verb +start+: the track start position 
\item \verb +end+: the track end position 
\item \verb +refName+: the genomic coordinates of the viewing window 
\end{itemize}
\end{itemize}

@@

\paragraph{Offset}
The left-margin offset requires dynamic syncing to ensure the scale between JBrowse and the heatmap remain aligned. The observable \verb +offsetCanvas+ is defined as a component of the \verb +PlotStateService+ class. The public variable \verb +offsetCanvasSource+ is defined as an instance of the \verb +BehaviourSubject+ class and holds all recorded values of the canvas offset. In the class constructor, \verb +offsetCanvas+ is initialised to continuously hold the latest value from \verb +offsetCanvasSource+. The canvas offset is set in the \verb +TreePlotComponent+ class, and is passed to \verb +offsetCanvasSource+ when the tree is redrawn.

The position of the JBrowse container is dynamically set in \verb+tersect-browser.component.ts+, with the \verb+style.margin-left.px+ set to \verb+offsetCanvas+, which holds thw pixel width of the phenetic tree. As a result, the position of the JBrowse container dynamically adjusts to different canvas states to ensure the scale bar between JBrowse and the heatmap remains aligned.

\paragraph{Horizontal Scroll}
Horizontal scroll inputs from the user are disabled inside the main JBrowse window through the implementation of a \verb+Plugin+ of the \verb+DisableScroll+ class in \verb+JbrowserWrapper.tsx+, which overrides the JBrowse configuration to prevent horizontal scrolling

@@

When a new dataset is added, the custom addition script automatically updates the names of tracks and assembly in the relevant files within the genome-browser extension. This flexibility allows the Genome Browser extension to be accessible by plant breeders working on many different datasets. TersectBrowser was initially validated with a human genome dataset, so it was important to maintain this flexibility of input in the new extensions.

\subsubsection{Future Improvements}
In addition to displaying the reference track, gene models track, and variants track, additional feature tracks, such as a multiple sequence alignment view, could be added to the genome browser.

Currently, the genome browser popup window is limited to opening a single window within the TersectBrowser+ webpage. A button to open the popup window in a new tab could be added, which would allow the user to open multiple detatched popup windows and view them simultaneously.

@@

\subsubsection{User Interface}
\paragraph{Feature Description}
\begin{itemize}
    \item The gene models track is displayed in the main JBrowse window. 
    \item In the popup window, the gene models track is displayed as the second track, underneath the reference sequence track and above the variant track.
\end{itemize}
